Data Analysis Project

### 2023-03-18 code/processing_code/processing.qmd
update manuscript and readme

```{r}
covid <- full_join(
  covid_symptom%>%
    select(cases.symptom.onset=cases, date=symptom.date), 
  covid_case%>%
    select(cases.reported=cases, date=report_date), 
  by = "date"
) %>% 
  full_join(
    covid_testing%>%
      rename(date=collection_date), 
    by="date"
  ) %>%
  select(date, cases.symptom.onset, cases.reported, pcr_tests, pcr_pos, pcr_pos_7dma, pcr_pos_14dma) 
  #full_join(cases.wrf, by = "date")

>>>>

```{r}
covid <- full_join(
  covid_symptom%>%
    select(cases.symptom.onset=cases, date=symptom.date), 
  covid_case%>%
    select(cases.reported=cases, date=report_date), 
  by = "date"
) %>% 
  full_join(
    covid_testing%>%
      rename(date=collection_date), 
    by="date"
  ) %>%
  select(date, cases.symptom.onset, cases.reported, pcr_tests, pcr_pos, pcr_pos_7dma, pcr_pos_14dma) %>%

  mutate(prop_pos = pcr_pos/pcr_tests)

### 2023-03-18 products/manuscript/Manuscript.qmd
update manuscript and readme

---
title: "Data Analysis Project"
subtitle: ""
author: Leah Lariscy
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

@@

Infection by Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2), causing COVID-19, is followed by the shedding of viral particles by multiple excretory functions, including stool and urine production.

Subsequently, these viral particles can be detected in wastewater influent via RNA extraction, followed by Reverse Transcriptase Quantitative Polymerase Chain Reaction (RT-qPCR) utilizing SARS-CoV-2 specific primers.

Wastewater-based detection methods have been utilized across the globe as an independent indicator of SARS-CoV-2 viral prevalence and, depending on community-specific factors, as a leading indicator of clinical case trends \[4\]. WBE methods may also become a primary tool in understanding transmission dynamics as case under reporting increases.

@@

24-hour composite samples collected twice weekly, stored at 4C until ready for extraction. 6 extraction replicates were performed for each collection date at treatment facility.

2019-nCoV N1 and N2 primer-probe sets were used for universal detection of SARS-CoV-2.

The concentration per reaction of each gene target was determined by the corresponding standard curve, then sample concentration in copies per liter was determined by Equation 1. Total daily copies per day per WRF was determined by Equation 2.

@@

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

@@

## Data aquisition

Wastewater SARS-CoV-2 data was obtained through weekly surveillance efforts utilizing RT-qPCR methods.

Water Reclamation Facility data was obtained through data share directly from the WRF representatives.

COVID data downloaded from Georgia DPH website.

@@

*qPCR data*

WWTP data

*DPH COVID data*

@@

Simple analyses: Wastewater and clinical data over time, linear regression model of wastewater vs clinical case data

Full analyses: Wastewater data as predictor for clinical case outcomes

Steps I took to build and tune model

@@

## Exploratory/Descriptive analysis

@@

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

@fig-result shows a scatterplot figure produced by one of the R scripts.

@@

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

Example @tbl-resulttable2 shows a summary of a linear model fit.

>>>>

---
title: "Data Analysis Project"
subtitle: ""
author: Leah Lariscy
date: "`r Sys.Date()`"
format:
  html:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(tidyverse)
library(skimr)
```

@@

Infection by Severe Acute Respiratory Syndrome Coronavirus 2 (SARS-CoV-2), causing COVID-19, is followed by the shedding of viral particles by multiple excretory functions, including stool and urine production. Subsequently, these viral particles can be detected in wastewater influent via RNA extraction, followed by Reverse Transcriptase Quantitative Polymerase Chain Reaction (RT-qPCR) utilizing SARS-CoV-2 specific primers. Wastewater-based detection methods have been utilized across the globe as an independent and parallel indicator of SARS-CoV-2 viral prevalence and, depending on community-specific factors, as a leading indicator of clinical case trends. WBE methods may also become a primary tool in understanding transmission dynamics as case under-reporting increases.

@@

### Wastewater SARS-CoV-2 surveillance data

24-hour composite wastewater samples were collected twice weekly from three treatment facilities in Clarke County and refrigerated until ready for extraction. RNA extractions were performed for each sample in replicates (n=6). Reverse Transcriptase Quantitative Polymerase Chain Reaction (RT-qPCR) and SARS-CoV-2 specific primer-probe pairs were used to amplify each RNA extraction replicate in technical replicates (n=3). The concentration per reaction of each gene target was determined by the amplification standard curve from positive controls corresponding to the targets.

Wastewater influent flow data, including total daily flow volume and total suspended solids rate, were collected for each corresponding sample date. This was used to estimate the viral load of SARS-CoV-2, also known as total daily copies.

### COVID-19 testing data

COVID-19 data sets were downloaded from the Georgia Department of Public Health website (accessed Jan 2023), of which a subset was selected which includes symptom onset, PCR test, and PCR positive test data for each date since the first case was detected.

@@

@@

## Data acquisition

**Wastewater SARS-CoV-2 viral load data** was obtained through weekly surveillance efforts utilizing RT-qPCR methods (Section 2.2).

**WWTP flow data** was obtained through data share directly from plant representatives.

**COVID-19 case data** was downloaded from Georgia Department of Public Health's COVID-19 Status Report Page in January 2023 \[x\].

@@

### RT-qPCR data

For each target, viral copies per microliter (cp/uL) of each PCR reaction was estimated by transforming the cycle threshold value (Ct) using the appropriate standard curve equation. For reactions with no Ct value (no detection), cp/uL was set to the theoretical limit of detection (LOD). From this, N1 and N2 data frames were combined and copies per liter (cp/L) of sample for both targets was estimated using Equation 1.

```{r, echo=FALSE, message=FALSE}

readRDS(here("data/processed_data/year2_cfx_clean.rds")) %>% glimpse()

```

### Treatment plant flow data

For each collection date, corresponding , were combined with the qPCR data. Viral load (total daily viral copies) was then estimated by multiplying concentration estimates by daily flow volume (Equation 2).

```{r, echo=FALSE, message=FALSE}
readRDS(here("data/processed_data/year2_cfx_plant_combined.rds")) %>% glimpse()
```

### COVID testing data

Data frames were combined into one and filtered to include only data pertaining to Clarke County. Proportion of positive tests per day was calculated by dividing PCR positive tests by total PCR tests.

```{r, echo=FALSE, message=FALSE}
readRDS(here("data/processed_data/all_covid_combined.rds")) %>% glimpse()
```

@@

### Exploratory analysis

Wastewater and clinical data over time, linear regression model of wastewater vs clinical case data

### Predicting clinical case levels with wastewater surveillance data

Wastewater data as predictor for clinical case outcomes

@@

## Descriptive analysis

@@

@@
